--- 
title: 'Report: Text mining'
author: "Claudio Previte and Ana -Maria Casian"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: TRUE
    toc_float:
      collapsed: true
      smooth_scroll: true
---



```{r message=FALSE, warning=FALSE}
library(here)
library(Hmisc)
library(dplyr)
library(DataExplorer)
library(ggplot2)
library(lattice)
library(inspectdf)
library(kableExtra)
library(explore)
library(DT)


```

# Loading data

```{r}
bands <- read.csv2(file = here('bands3.csv'), sep = ';',na.strings = "?")
```


# First visaulization
```{r}
datatable(bands)


bands %>% 
  introduce() %>%
  t()%>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed","bordered"),full_width = F, position = "center")%>%
  footnote(symbol = strwrap("Short description of the data"))

```


#Check the data structure

```{r, message=FALSE, warning=FALSE, include= FALSE}
str(bands) #this part will not apear in the report, but we should comment about it. This line of code helped us to decide wich data to transform in the first part

```


#Data Transformation

```{r message=FALSE, warning=FALSE}

bands <- as.data.frame(lapply(bands,function(x) 
  if(is.factor(x)) factor(toupper(x)) 
  else(x))) # uppercase for all the factor values

#there is a warning message!!!! I took it out for now, but to review it and understand it


cols = c(21:39)  
bands[,cols] = apply(bands[,cols], 2, function(x) as.numeric(as.character(x))) #change class to numeric for num row

bands <- transform(bands, date = as.Date(as.character(date), "%Y%m%d"))



```

#Double-cheking the data structure is correct
```{r , message=FALSE, warning=FALSE, include= FALSE}

#make sure that the variables are well defined
str(bands)

```




```{r}



library(VIM)
aggr(bands[, which(sapply(bands, function(x)all(any(is.na(x)))) == TRUE)], sortVar=TRUE, cex.lab=0.8, cex.axis=0.9,  numbers = TRUE, combined = TRUE,  oma = c(8,1,1,1))



```


```{r}
library(dplyr)

bands <- bands[-c(486:540),]

bands$location = as.character(bands$location)
bands$location[is.na(bands$location)] <- "ana"
bands$location = as.factor(bands$location)

bands <-
  bands[, which(
    !names(bands) %in% c(
      "date",
      "ink_color",
      "customer"
      
    )
  )]


Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

bands<- as.data.frame( lapply(bands, function(x) {
    if(is.factor(x)) replace(x, is.na(x), Mode(na.omit(x)))
    else if(is.numeric(x)) replace(x, is.na(x), mean(x, na.rm=TRUE))
    else x
}))


library(Boruta)
boruta_test <- Boruta(band_type~., data = bands, doTrace = 2)



kable(boruta_test$finalDecision)
info_boruta = as.data.frame(boruta_test$finalDecision)
colnames(info_boruta) <- c("decision")

plot(boruta_test, xlab = "", xaxt = "n")
lz <- lapply(1:ncol(boruta_test$ImpHistory), function(i)
  boruta_test$ImpHistory[is.finite(boruta_test$ImpHistory[, i]), i])
names(lz) <- colnames(boruta_test$ImpHistory)
Labels <- sort(sapply(lz, median))
axis(
  side = 1,
  las = 2,
  labels = names(Labels),
  at = 1:ncol(boruta_test$ImpHistory),
  cex.axis = 0.8
)

names(lz) <- colnames(boruta_test$ImpHistory)

final_decision= as.data.frame(boruta_test$finalDecision)

setDT(final_decision, keep.rownames = TRUE)[]
colnames(final_decision) <- c("variable","decision")

boruta.data <- final_decision %>%
  filter(final_decision$decision == 'Confirmed' | final_decision$decision == 'Tentative')

variables_to_keep <- boruta.data$variable



```

# Deleting useless variables

To write why we consider the useless. 
```{r}



bands <-
  bands[, which(
    names(bands) %in% variables_to_keep
  )]


```



#Outliers. We have generated the boxplot and the summary for all the variables and examinated the outliers. Given that the outliers were not showing any significant values, we have decided that there are no error meassurements and neither error entries in the data

```{r, message=FALSE, warning=FALSE, include= FALSE}

numeric_column <- names(select_if(bands, is.numeric))

for (i in numeric_column) {
  print(summary(bands[i]))
  print (boxplot(bands[i])$out)
}

```





#Summary bands

```{r}
# summary for 

bands %>% 
  explore::describe() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed","bordered"),full_width = F, position = "center")

bands %>%
  summary() %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover","condensed","bordered"),full_width = F, position = "center")

plot_intro(bands)
```




```{r fig.height= 12}


plot_histogram(bands)

plot_density(bands)

bands %>% explore_all(target = band_type)

plot_boxplot(bands, by= 'band_type',  ncol = 2, title = "Side-by-side boxplots")

plot_correlation(bands, type= 'c', cor_args = list( 'use' = 'complete.obs'))

bands %>% 
  select(numeric_column, band_type) %>%
  explore_all(target = band_type)

#ggpairs(bands[,-40], ggplot2::aes(colour=band_type))

# split data in 2
bands.band <- filter(bands, bands$band_type == 'BAND')

bands.noband <- filter(bands, bands$band_type == 'NOBAND')

# summary(bands.band)
# summary(bands.noband)

# managing missing values (19 rows with NAs)

na <- inspect_na(bands)
show_plot(na, col_palette=2)
```


```{r}
bands[,c(12:28)]  %>% 
  correlate() %>% network_plot(min_cor = .1)

```


```{r bands.ct}
require(rpart)
library(rpart.plot)
set.seed(123456)
bands.ct <- rpart(band_type ~ ., method = "class", data = bands, control = rpart.control(minsplit = 4, 
    cp = 1e-05), model = TRUE)

summary(bands.ct)

rpart.plot(bands.ct, box.palette="RdBu", nn=FALSE)

# par(mar = c(0.5, 1, 0.5, 1))
# plot(bands.ct, uniform = TRUE)
# text(bands.ct, use.n = TRUE, all = TRUE, cex = 0.6)


```

```{r bands.par}

require(lattice)

trellis.par.set(col.whitebg())
plot.bands.par <- xyplot(roller_durometer ~ viscosity, bands, groups = band_type, pch = 16, 
    col = c("red", "green"), panel = function(x, y, groups, ...) {
        panel.superpose(x, y, groups, ...)
        panel.abline(v = 2.45, lty = 2)
        panel.segments(2.45, 1.75, max(x) * 2, 1.75, lty = 2)
        panel.segments(4.95, min(y) * -2, 4.95, 1.75, lty = 2)
        panel.segments(2.45, 1.65, 4.95, 1.65, lty = 2)
        panel.segments(4.95, 1.55, max(x) * 2, 1.55, lty = 2)
    }, key = list(columns = 3, col = c("red", "green"), text = list(c("BAND", 
        "NOBAND"))))
print(plot.bands.par)

printcp(bands.ct)

par(pty = "s")
plotcp(bands.ct)

# pruning trees
par(pty = "s")
with(bands.ct, plot(cptable[, 3], xlab = "Tree Number", ylab = "Resubstitution Error (R)", 
    type = "b"))

par(pty = "s")
with(bands.ct, plot(cptable[, 4], xlab = "Tree Number", ylab = "Cross-Validated Error (R(cv))", 
    type = "b"))

par(pty = "s")
plotcp(bands.ct)
with(bands.ct, {
    lines(cptable[, 2] + 1, cptable[, 3], type = "b", col = "red")
    legend(3, 1, c("Resub. Error", "CV Error", "min(CV Error)+1SE"), lty = c(1, 
        1, 2), col = c("red", "black", "black"), bty = "n")
})

bands.prune <- prune(bands.ct, cp = 0.1)
summary(bands.prune)

library(knitr)
bands.pred <- predict(bands.prune, type = "class")
table(bands.pred, bands$band_type)
```

```{r neural network model}

require(nnet)

bands.net <- nnet(band_type ~ ., data = bands, size = 2, rang = 0.1, decay = 5e-04, 
    Hess = T, maxit = 200)

eigen(bands.net$Hess, T)$values

library(NeuralNetTools)
par(mar = numeric(4), family = "serif")
plotnet(bands.net, pos_col = "darkgreen", neg_col = "darkblue")
```

